A decomposition based on constrained clustering algorithms for JSP\thanks{Supported by organization x.}

### 2021-09-27 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Modifying main issues

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{changepage}
\usepackage{cite}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{A decomposition based on constrained clustering algorithms for job shop scheduling problems\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{1}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{1}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{M. El-Kholany et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
Technische Universität Graz, Graz, Austria\\
\email{mgebser@ist.tugraz.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Scheduling decision in production planning plays a significant role in the manufacturing industry. A good schedule may lead to solve production problems and improve productivity. Job-shop Scheduling Problem (JSP) is one of the most crucial problems in production systems. JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine while completing all tasks as early as possible. One of the most effective approaches studied by many researchers for solving the JSP is decomposition. This work investigates applying the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies. This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window. The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine. These features are extracted either from the problem itself or from solutions obtained by other heuristics. The second phase is to develop a constrained clustering algorithm to assign each data object (an operation) into a cluster. We solved the problem using the Answer Set Programming model, and we tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases. The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest starting Time} have a significant impact on the solution quality.

@@

Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}. 
%\comment{Are there any citations for the claim above?}
Scheduling is one of the most complex problems in the industry. Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP). The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard \cite{baker1974introduction,lenstra1979computational}. Therefore, it is very difficult to reach an optimal solution in a reasonable time, even for small instances. However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}.

From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine. Our study aims to  find the most appropriate sequence of operations to optimize the typical performance indicator (makespan). Several approaches have been studied to solve JSP, and one of the most effective approaches used is decomposition \cite{zhang2010hybrid,zhai2014decomposition}. The main idea of the decomposition is to split the problem into a series of sub-problems (windows) and solve each of them separately in a sequence, and then obtain the solution of the original problem.\\

This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects. For the JSP, an instance consists of a set of jobs to be processed on machines. The number of jobs is $n$, and the number of machines is $m$. Each job contains $m$ operations with the operation precedence constraints. Each operation should be executed on machine $m$ for a certain amount of time. Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}.\\

@@

    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a time window. It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
\end{enumerate}

\section{Features extraction}
In this section, we will show which features have been extracted and how did we extract them. Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position. Most of the researchers converted the attributes into qualitative data. As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems. However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
We will show the features that we considered:
\subsection{\textit{Process}}
\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain. The \textit{PT} represents the execution time of a particular operation in a machine. However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.

@@

\subsection{\textit{Time Length of a Job}(TLJ)}
\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.

@@

\subsection{\textit{Waiting Time} (WT)}
Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times. \textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free.

@@

In this section, we will show the implementation of our proposed model and how it works. The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem. We aim to decompose the problem into time windows because this technique proved its effectiveness for solving the scheduling problems. Several ways have been presented during the last decades to decompose the problem  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\

The proposed model aims to apply clustering methods where each cluster will represent a time window, and each operation is a data object. The clustering algorithms are used to group a set of data objects into one cluster. It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects. The similar objects will be in one cluster, and the dissimilar are in different clusters. The difference between the basic clustering and our proposed is the ordering of the clusters. In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar. However, in our case, the cluster number is meaningful because the data objects grouped in cluster $i$ should be scheduled before other data objects in cluster $i+1$. More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s). Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster. So that, we will have two types of constraints:
\begin{enumerate}
\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
\item The number of operations assigned to each cluster should be close to each other.
\end{enumerate}

Fig.~\ref{fig1} shows a flowchart of the proposed algorithm. As shown in the flowchart, the number of clusters is given. We calculate the cluster capacity by dividing the total number of operations by the number of clusters. In the next step, the clustering process started by generating the centroids, where the number of centroids is the number of clusters. In general, the centroids are generated randomly as well as in our model. However, we selected some operations as centroids. Although the operations are selected randomly, we followed a rule to select the centroids. For example, let us assume that a job consists of 15 operations, and three centroids should be generated. The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations. We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs. After generating the centroids, we perform a double-check for exceeding the number of clusters and the cluster capacity, respectively. After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster. In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated. If the maximum number of operations is reached, we will go to the next cluster. After assigning all operations into clusters, the algorithm terminates.

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{Flow_chart_9.png}
  \caption{Flowchart of the proposed Clustering Algorithm.}
  \label{fig1}
\end{figure}

@@

@@

We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments. The following Table~\ref{tab1} shows the selected features, and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}. The following table shows the results of our experiments. The first column represents the instances we tested, and the next three columns show the makespan values obtained by FIFO, MTWR, and EST heuristics. From the fifth column, we show the makespan values obtained by combining the features mentioned above. Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics. However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, \textbf{F14} and \textbf{F15}) have a good impact in the cost function in the most cases. Especially, combinations \textbf{F6} and \textbf{F15} provided good results. It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}. Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand. More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics.
\begin{table}[h!]
\setlength{\tabcolsep}{4.0pt}
  \begin{center}
    \caption{The performance of selected features.}
    \label{tab1}
 \begin{adjustwidth}{-3.7cm}{}
    \begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
      \hline
      TA51 &3549  &3364  &3833  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
      TA52 &3339  &3304  &3594  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
      TA53 &3160  &3168  &3443  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
      TA54 &3218  &3494  &3439  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
      TA55 &3291  &3237  &3784  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
      TA56 &3325  &3287  &3552  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
      TA57 &3654  &3633  &3681  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
      TA58 &3299  &3591  &3764  &3352  &3412  &3370  &3509  &3707  &\textbf{3324}  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
      TA59 &3344  &3394  &3661  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
      TA60 &3129  &3257  &3399  &3483  &\textbf{3315}  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
    \end{tabular}
\end{adjustwidth}
  \end{center}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{helvet,times,courier}
\usepackage{graphicx}
\usepackage{changepage}
\usepackage{cite}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{caption} 
\captionsetup[table]{skip=10pt}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{A decomposition based on constrained clustering algorithms for JSP\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
Martin Gebser\inst{1,2}\orcidID{0000--0000-0000-0000}}
%
\authorrunning{M. El-Kholany et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
Technische Universität Graz, Graz, Austria\\
\email{mgebser@ist.tugraz.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Scheduling decision in production planning plays a significant role in the manufacturing industry. A good schedule may lead to solve production problems and improve productivity. Job-shop Scheduling Problem (JSP) is one of the most crucial problems in production systems. JSP deals with allocating tasks on resources and determining the sequence of the operations on each machine while completing all tasks as early as possible. One of the most effective approaches studied by many researchers for solving the JSP is decomposition. This work investigates applying the decomposition strategy to split the problem into a series of sub-problems using data mining methodologies. This study proposes a constrained clustering algorithm to group the data objects (operations) into clusters, where each cluster is a Time Window (TW). The decomposition process depends on two main phases; the first phase is to extract features to predict the sequence of the operations on each machine. These features are extracted either from the problem itself or from solutions obtained by other heuristics. The second phase is to develop a constrained clustering algorithm to assign each data object (an operation) into a cluster. We solved the problem using the Answer Set Programming model, and tested our proposed algorithm on a set of benchmark instances, and the results showed that our proposed outperformed other heuristics in most cases. The results also showed that some features like \textit{Remaining Processing Time}, \textit{Machine Load} and \textit{Earliest Starting Time} have a significant impact on the solution quality.

@@

Clustering has been widely applied in several disciplines; one is production-planning systems~\cite{nananukul2013clustering,koskosidis1992clustering}, more specifically scheduling \cite{yashar2013multi,tong2016research}. 
%\comment{Are there any citations for the claim above?}
Scheduling is one of the most complex problems in the industry. Scheduling of operations involves resource allocation over a period of time to perform a series of tasks, where one of the most critical problems is the Job-shop Scheduling Problem (JSP). The JSP has been known as a complex and combinatorial optimization problem, and it has been proven to be NP-hard \cite{baker1974introduction,lenstra1979computational}. Therefore, it is very difficult to reach an optimal solution in a reasonable time, even for small instances. However, from a practical perspective, the JSP scale is large, where the number of operations can be up to 10,000s in some workshops \cite{zhang2010hybrid}. \\

This work intends to apply the clustering techniques to decompose the problem into sub-problems, where the number of clusters corresponds to the number of sub-problems (windows), and the operations correspond to data objects. For the JSP, an instance consists of a set of jobs to be processed on machines. The number of jobs is $n$, and the number of machines is $m$. Each job contains $m$ operations with the operation precedence constraints. Each operation should be executed on machine $m$ for a certain amount of time. Since the problem at hand has precedence constraints, it is required to apply a constrained clustering approach for the decomposition \cite{wagstaff2001constrained}. Although many studies proposed different constrained clustering methods in \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}. However, most of those checked whether two data objects must be assigned into the same cluster. For the JSP, it is essential to consider the sequencing between the clusters to handle the precedence constraints. More specifically, our proposed method will not permit to assign an operation into a particular cluster if the predecessor of that operation is not yet assigned to the same or previous cluster(s). \\

@@

    \item The second phase aims to develop, implement and test a proposed clustering algorithm that will split the problem into sub-problems where each operation is a data object, and each cluster will represent a Time Window (TW). It will consider the precedence constraints between the operations and balancing between the clusters according to the number of operations per cluster.
\end{enumerate}

The paper is organized as follows. The following Section describes the Job-shop Scheduling Problem in detail and shows an illustrative example of input data. Section $3$ shows which features we extracted and how we extracted them, followed by Section $4$, which presents the proposed algorithm. Section $5$ presents an extensive computational study on benchmark instances, comparing our results with the others obtained by some heuristics. A relevant literature review on the subject is given in Section $6$ followed by the conclusion and future work.

\section{Problem Formulation}
The Job-shop Scheduling Problem (JSP) is one of the well-known hardest combinatorial optimization problems \cite{baker1974introduction,lenstra1979computational,garey1976complexity}. The JSP is given a set $J = \{ J_1, J_2,...,J_n \}$  of jobs and a set $M = \{ M_1, M_2,...,M_m \}$ of Machines. Each job $J_i$ has a series $  O_{i,1}, O_{i,2},...,O_{i,m} $ of operations that must be processed in a predefined order. Each machine $M_i$ executes an operation of each job with a known and fixed processing time. The main objective of the JSP is to minimize the total completion time for all jobs (makespan). \\ The JSP assumptions are listed as follows:

\begin{enumerate}
	\item Once an operation is started, it must be completed without interruption.
	\item Each job visits a machine once.
	\item A machine cannot perform more than one operation at a time.
	\item A machine is not broken.
	\item All jobs and machines are available at time 0.
\end{enumerate}
It is hard to reach a solution close to the optimal in a reasonable time, even if the number of operations to be scheduled is low. In real-life applications, the number of operations that are to be scheduled is up to $10,000$ \cite{zhang2010hybrid}, which makes the problem more complicated. The JSP has attracted the interest of many researchers in the last decades to find effective methods to solve this problem. One of these methods is decomposition. The decomposition idea is to split the problem into sub-problems, solve them separately, and then integrate those solutions to solve the whole problem. Many decomposition strategies have been proposed for solving the JSP  \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}. However, there is no guarantee of which method is the best because the performance varies according to the problem. \\

A problem data can be organized in a table, where rows correspond to operations and columns refer to the name of the operation, the machine assigned to an operation, and the processing time of a particular operation. Table~\ref{tab1} shows the information of a simple Job-shop Scheduling Problem which consists of $3$ jobs and $3$ machines. For instance, in the third row of the table, $O_{1,3}$ is the third operation of the first job that will be executed by machine $1$ with processing time $5$.

\begin{table}[h!]
\fontsize{12}{9}
\setlength{\tabcolsep}{10.0pt}
\centering
  \begin{center}
    \caption{An example of a Job-shop Scheduling Problem.}
    \label{tab1}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c c c}
      \textbf{Operation} & \textbf{Machine} & \textbf{Procession Time} \\
      \hline
								\\
      $O_{1,1}$  & 2  &  9  	\\
      $O_{1,2}$  & 3  &  3  	\\
      $O_{1,3}$  & 1  &  5  	\\
								\\
      $O_{2,1}$  & 3  & 4  	\\
      $O_{2,2}$  & 2  & 6  	\\
      $O_{2,3}$  & 1  & 2  	\\
								\\
      $O_{3,1}$  & 1  & 4  	\\
      $O_{3,2}$  & 3  & 3  	\\
      $O_{3,3}$  & 2  & 5  	\\
    \end{tabular}

%\end{adjustwidth}
  \end{center}

\end{table}
In our example in Table~\ref{tab1}, there are $9$ operations and $3$ machines. We aim to split them efficiently into two parts and schedule them separately without violating the precedence constraint. More specifically, we should not assign an operation in the first TW, and its predecessor is assigned to the following TW(s). Table~\ref{tab2} shows how the output of the proposed method should look. As shown in the Table~\ref{tab2}, operations $\{ O_{1,1}, O_{1,2}, O_{2,1}, O_{2,2}, O_{3,1} \}$ are assigned to the first TW and the rest to the second TW. It means that the first TW will be optimized and then the upcoming TW(s). In this study, we investigate applying a constrained clustering algorithm to decompose the JSP, where we extract some features from the problem itself and the solutions of other heuristics. These features have been used for the splitting process, and we present them in detail in the following section.

\begin{table}[h!]
\fontsize{12}{9}
\setlength{\tabcolsep}{10.0pt}
\centering
  \begin{center}
    \caption{Time Window Assignment.}
    \label{tab2}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c }
      \textbf{Operation} & \textbf{Time Window}  \\
      \hline
							\\
      $O_{1,1}$  & 1    	\\
      $O_{1,2}$  & 1    	\\
      $O_{2,1}$  & 1    	\\
      $O_{2,2}$  & 1   	\\
      $O_{3,1}$  & 1		\\
					    		\\
      $O_{1,3}$  & 2  	\\
      $O_{2,3}$  & 2		\\
      $O_{3,2}$  & 2    	\\
      $O_{3,3}$  & 2    	\\
    \end{tabular}
%\end{adjustwidth}
  \end{center}

\section{Features extraction}
In this section, we show which features have been extracted and how did we extract them. Based on the previous work, \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data}, the following features have been identified: priority, processing time, remaining processing time, machine load, and route position. Most of the researchers converted the attributes into qualitative data. As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems. However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
We present the features that we considered:
\subsection{\textit{Process}}
\textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are attributes dependent on the problem domain. The \textit{PT} represents the execution time of a particular operation in a machine. However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job.

@@

\subsection{\textit{Time Length of a Job} (TLJ)}
\textit{TLJ} is an attribute that describes the total processing time needed to complete a particular job. It is the summation of processing time for all operations of that job.

@@

\subsection{\textit{Waiting Time} (WT)}
Since we used three different heuristics to mine information from the obtained solution, we will have three waiting times. \textit{WT} attribute is the difference between the starting time of a particular operation and the maximum value between the end time of its predecessor and when the machine became free. \\

The following Table~\ref{tab3} shows a sample of the features that we extracted for the decomposition process. Each row in the table represents an operation, and it contains its value according to each feature. The second column shows the \textit{Remain Processing Time (RPT)} of a job. For instance, the first job $J_1$ consists of $3$ operations with a total processing time of $17$. In the beginning, the \textit{Remain Processing Time} of $O_{1,1}$ is $17$ because no operations are executed. After finishing the first operation of job $1$, the \textit{RPT} became $8$ because the processing time of the completed operation $O_{1,1}$ is excluded. When the second operation of job $1$ is completed, the \textit{RPT} of that job is $5$. \\

The third column represents the \textit{Earliest Starting Time} which we described earlier. The last column is the \textit{Machine Load} which is the summation of the processing time for all operations assigned to a machine. For example, Machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$, the load of machine $3$ at time $0$ is 10. We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}. After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.

\begin{table}[h!]
\fontsize{12}{9}
\setlength{\tabcolsep}{10.0pt}
\centering
  \begin{center}
  \caption{Time Window Assignment.}
  \label{tab3}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c  c  c}
      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML}\\
      \hline
										\\
      $O_{1,1}$  & 17 & 0   & 20	\\
      $O_{1,2}$  & 8  & 9   & 3		\\
      $O_{1,3}$  & 5  & 12 & 5		\\
										\\
      $O_{2,1}$  & 12 & 0   & 10	\\
      $O_{2,2}$  & 8  & 4   & 11	\\
      $O_{2,3}$  & 2  & 10 & 7		\\
					    					\\
      $O_{3,1}$  & 12 & 0  & 11	\\
      $O_{3,2}$  & 8  & 4  & 6		\\
      $O_{3,3}$  & 5  & 7  & 5		\\
    \end{tabular}
%\end{adjustwidth}

  \end{center} 
\end{table}

@@

In this section, we will show the implementation of our proposed model and how it works. The main idea of the proposed model is to split the problem into sub-problems, solve them sequentially and integrate the solutions to have the solution of the main problem. We aim to decompose the problem into TWs because this technique proved its effectiveness for solving the scheduling problems \cite{zhang2010hybrid,zhai2014decomposition}. Several ways have been presented during the last decades to decompose the problem \cite{zhai2014decomposition,singer2001decomposition,ovacik2012decomposition,uzsoy2000performance}.\\

The proposed model aims to apply clustering methods where each cluster will represent a TW, and each operation is a data object. The clustering algorithms are used to group a set of data objects into one cluster. It depends on measuring the similarity/dissimilarity between the data objects according to the features of these data objects. The similar objects will be in one cluster, and the dissimilar are in different clusters. The difference between the basic clustering and our proposed is the ordering of the clusters. In general, the cluster number does not have meaning; it just means that the data objects grouped in a particular cluster are similar. However, in our case, the cluster number is meaningful because the data objects grouped in a particular cluster $i$ should be scheduled before other data objects in the successor cluster(s). More specifically, it is not allowed to insert a data object (\textit{an operation}) in a particular cluster, and the predecessor of that operation is not assigned yet to the same cluster or the preceding cluster(s). Since we will solve each window separately, we should consider balancing the clusters concerning the number of operations assigned to each cluster. So that, we will have two types of constraints:
\begin{enumerate}
\item The predecessor(s) of an operation must be inserted into the same or previous cluster(s).
\item The number of operations assigned to each cluster should be close to each other.
\end{enumerate}

Algorithm.~\ref{alg} shows the pseudocode of the proposed algorithm. As shown in the pseudocode, the number of clusters is given. We calculate the cluster capacity by dividing the total number of operations by the number of clusters. In the next step, the clustering process starts by generating the centroids, where the number of centroids is the number of clusters. In general, the centroids are generated randomly as well as in our model. However, we selected some operations as centroids. Although the operations are selected randomly, we followed a rule to select the centroids. For example, let us assume that a job consists of 15 operations, and the number of clusters is $3$. The first centroid will be an operation of rank 1 - 5, the second centroid will be 6 - 10, and the third will be from the last five operations. We followed this strategy to make them far from each other because the operations grouped in the first clusters should be at the beginning of their jobs. After generating the centroids, we check for exceeding the number of clusters and the cluster capacity, respectively. After that, we calculate the distance between each operation and the centroid of the current cluster using Euclidean distance and assign the nearest operation to that cluster. In order to satisfy the precedence constraint, we check if the predecessor is not assigned to a cluster, it will be assigned to the current cluster, and the centroid is updated. If the maximum number of operations is reached, we will go to the next cluster. After assigning all operations into clusters, the algorithm terminates.

\begin{algorithm}
\caption{Proposed Clustering Algorithm}\label{alg}
\begin{algorithmic}
\Require $num\_clus$
%\Require $n \geq 0$
%\Ensure $y = x^n$
\State $count\_clus \gets 0$

\State $clus\_capacity \gets \ceil*{\frac{num\_oper}{num\_clus}}$
\State Generate centroids
\While{$count\_clus < num\_clus$}
    \State $count\_cap \gets 0$
    \State $curr\_clus \gets count\_clus + 1$
	\While{$count\_cap < clus\_capacity$}
          \State Calculate distance between data objects and a centroid   \Comment{Using Euclidean distance}
          \State $selected\_data\_obj \gets nearest\_data\_obj$
          %\State $selected\_data\_obj \in curr\_clus$
          \State $curr\_clus \gets selected\_data\_obj$    \Comment{Assigning an operation to a TW}

          \If{$pred\_selected\_data\_obj \notin clus$}     \Comment{Satisfying the precedence constraint}
              \State $curr\_clus \gets pred\_selected\_data\_obj$
          \EndIf
          \State $update\_centroid$
          \State $count\_cap \gets count\_cap + 1 $
	\EndWhile
    \State $count\_clus \gets count\_clus + 1 $
\EndWhile
\end{algorithmic}
\end{algorithm}

\iffalse
\begin{figure}[h!]
  \includegraphics[width=\linewidth]{Flow_chart_9.png}
  \caption{Flowchart of the proposed Clustering Algorithm.}
  \label{fig1}
\end{figure}
\fi

%\newpage

@@

\fontsize{9pt}{10pt}

@@

We neglected the rank of an operation and the time length of a job because they had negative impact results from the preliminary experiments. The following Table~\ref{tab4} shows the selected features and the performance of each combination for solving benchmark instances of 50 jobs and 15 machines \cite{taillard1993benchmarks}. The following table shows the results of our experiments. The first column represents the instances we tested, and the next four columns show the makespan values obtained by FIFO, MTWR, and (EST with single and multi-shot) heuristics. From the sixth column, we show the makespan values obtained by combining the features mentioned above. Our experiments found that the first five combinations do not provide good solutions compared to the other heuristics. However, features (\textbf{F6}, \textbf{F8}, \textbf{F11}, \textbf{F12}, and \textbf{F15}) have a good impact in the cost function in the most cases. Especially, combinations \textbf{F6} and \textbf{F15} provided good results. It means that the most important features extracted from the problem itself and have a positive impact on the performance are \textit{Remaining Processing Time, Machine Load and Earliest Starting Time}. Also, The information obtained from the solutions of the other heuristics like (\textit{Starting Time} and \textit{Waiting Time}) led us to get better decomposition for the problem at hand. More specifically, combining \textit{Remaining Processing Time} with \textit{Machine Load} provides a good decomposition with/without \textit{Waiting Time} attributes obtained by the other heuristics. In a few instances, our proposed method did not outperform other heuristics, and the average makespan of \textbf{F6} was better than the average makespan obtained by the other heuristics and close to the FIFO makespan. The following section illustrates the difference between our proposed algorithm and the methods implemented by other researchers.

%\begin{table}
\begin{sidewaystable}
\setlength{\tabcolsep}{5.5pt}
  \begin{center}
    \caption{The performance of selected features.}
    \label{tab4}
%\rotatebox{90}{
 \begin{adjustwidth}{-3.5cm}{}
    \begin{tabular}{c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c | c}

\hline
      \textbf{Instances} & \textbf{FIFO} & \textbf{MTWR} & \textbf{Single-shot} & \textbf{EST} & \textbf{F1} & \textbf{F2} & \textbf{F3} & \textbf{F4} & \textbf{F5} & \textbf{F6} & \textbf{F7} & \textbf{F8} & \textbf{F9} & \textbf{F10} & \textbf{F11} & \textbf{F12} & \textbf{F13} & \textbf{F14} & \textbf{F15}\\
      \hline
      TA51 &3549  &3364  &3632  &3506  &3506  &3362  &3324  &3360  &3346  &\textbf{3294}  &3328  &3588  &3588  &3568  &3514  &3682  &3624  &3554  &3509 \\
      TA52 &3339  &3304  &3615  &3773  &3277  &3318  &3330  &3286  &3243  &3275  &3226  &3397  &3373  &3533  &3256  &3494  &3505  &3373  &\textbf{3184} \\
      TA53 &3160  &3168  &3481  &3478  &3382  &3585  &3347  &3543  &3517  &3250  &3338  &3251  &3522  &3373  &3585  &\textbf{3143}  &3427  &3524  &3380 \\
      TA54 &3218  &3494  &3462  &3497  &3414  &3425  &3425  &3503  &3397  &3265  &3353  &3563  &3443  &3386  &3352  &3270  &3236  &3232  &3289 \\
      TA55 &3291  &3237  &3552  &3482  &3308  &3441  &3424  &3319  &3355  &3386  &3301  &3498  &3385  &3530  &3365  &3309  &3370  &3385  &\textbf{3218} \\
      TA56 &3325  &3287  &3610  &3605  &3353  &3380  &3548  &3270  &3383  &3366  &3564  &\textbf{3229}  &3306  &3362  &3344  &3332  &3338  &3483  &3389 \\
      TA57 &3654  &3633  &3778  &3753  &3605  &3573  &3601  &3583  &3650  &\textbf{3480}  &3592  &3621  &3543  &3608  &3670  &3632  &3749  &3573  &\textbf{3480} \\
      TA58 &3299  &3591  &3718  &3731  &3352  &3412  &3370  &3509  &3707  &3324  &3496  &3517 &3716 &3644  &3611  &3471  &3437  &3365   &3481 \\
      TA59 &3344  &3394  &3613  &3398  &3453  &3416  &3301  &3339  &3508  &3430  &3352  &3258 &3381  &3402  &\textbf{3145}  &3250  &3253  &3284  &3290 \\
      TA60 &3129  &3257  &3550  &3247  &3483  &3315  &3617  &3563  &3591  &3327  &3399  &3374  &3526  &3637  &3425  &3676  &3659  &3767  &3569 \\
\hline
 Average  &3331  &3373  &3601  &3547  &3413  &3423  &3429  &3428  &3470  &3340  &3395  &3430  &3475  &3504  &3427  &3426  &3460  &3454  &3379 \\
\hline
    \end{tabular}
%}
\end{adjustwidth}
  \end{center}
\end{sidewaystable}
%\end{table}

From the literature, we have found that most of the studies which used data mining techniques aimed to extract some features either from the problem instances or solutions obtained by some heuristics to build a decision-rules to order the operations assigned to a machine.\\
